### 2011-04-13 paper/methods.tex
analysis

\section{Methods}
\newcommand\few{few}
We ran the system with several hundred different combinations of features.
We used two approaches to identifying the best feature combinations, and then
we did more detailed analysis on the top \few systems.

\subsection{Identification of the best feature combinations}
We identified optimal feature combinations through backward stepwise simple linear regression
and through the creation and visual analysis of charts. The former has the
advantage of discovering relationships and testing them more quantitatively
but has the disadvantage of making assumptions about normality, linearity and equal variance.
The latter has the advantage of elucidating more complex and higher-order relationships

\subsubsection{Stepwise regression}
We ran backward stepwise regression with the F-measure from the coarse-grained scores
regressed on predictors representing the feature combination.
The saturated model contained the following predictors and all interactions
\begin{itemize}
\item Classifier (naive Bayes or decision list)
\item Colocation window length (0--8)
\item Cooccurrence handling (on or off)
\item Base word handling (on or off)
\item Bootstrap iterations (0--4)
\item Dependency parsing (on or off)
\end{itemize}
The predictor combination yeilding the lowest value of the Akaike information criterion
included the following parameters.
\begin{itemize}
\item Intercept
\item Classifier
\item Colocation window size
\item Coocurrence
\item Base word handling
\item Classifier$\times$colocation window size interaction
\item Classifier$\times$cooccurrence interaction
\end{itemize}
This does not include bootstrap iterations, indicating that changes
in bootstrap iterations does not substantially change the F-measure.

%Discuss the coefficients in the regression table
                         Estimate Std. Error t value Pr(>|t|)    
(Intercept)               0.572768   0.006419  89.229  < 2e-16 ***
classifiert               0.050648   0.015703   3.225  0.00162 ** 
colocation                0.003961   0.001146   3.455  0.00076 ***
cooccurrence              0.025666   0.005986   4.288 3.65e-05 ***
base_word                 0.011615   0.005599   2.074  0.04017 *  
classifiert:colocation   -0.013670   0.005475  -2.497  0.01388 *  
classifiert:cooccurrence -0.025294   0.016931  -1.494  0.13780

%lm(formula = f(foo) ~ classifier + colocation + cooccurrence + 
%    base_word + classifier:colocation + classifier:cooccurrence, 
%    data = foo)

\subsubsection{Graphical methods}

\subsection{Characterization of the top feature combinations}